Hunting experience shapes individual foraging specialization and predator-prey interactions in an online videogame

### 2023-06-01 manuscript/Chapter2-manuscript.Rmd
added some elements to the hypotheses and updated the methods

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [@Griffen.etal2012;@Michalko.Pekar2016;@Moran.etal2017;@Michalko.etal2021]. Indeed, predator populations often comprise assemblages of individuals specializing on different foraging tactics, with consequences for habitat use, functional responses, prey choice, and foraging rate [@Kobler.etal2009;@toscanoTraitmediatedFunctionalResponses2014;@Patrick.Weimerskirch2014;@Matsumura.Miyatake2022a]. Despite the clear evidence that individual specialization is important for predator-prey interactions, its underlying mechanisms remain surprisingly unclear. One hypothesis suggests that reciprocal interactions between predators and prey could drive individual foraging specialization [@Araujo.etal2011;@toscanoPersonalityForagingBehavior2016;@DeLisle.etal2022]. For instance, predators often hunt using techniques that are fine-tuned to the type of prey that they encounter [@Estes.etal2003;@Woo.etal2008;@Courbin.etal2018], and their refinement is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@Heithaus.etal2018]. Thus, the capacity of predators to effectively use hunting tactics should be contingent on periods of extensive practice (i.e. experience), which may result in different foraging specializations.

Theory predicts that foraging specialization may emerge via learning and expertise acquisition [@Tinker.etal2009;@Dukas2019]. @Dukas2019 defines expertise as the characteristics, skills, and knowledge allowing individuals with extensive experience to outperform novices on complex tasks. It is described as the relationship between success and experience, where extensive practice optimizes success when individuals reach an advanced level of experience. Empirical studies on human and non-human hunters show that experience optimizes foraging efficiency (e.g. search and handling times, return rates) potentially via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@MacDonald2007;@Reid.etal2010;@Wilson-Rankin2015]. This body of work suggests that such optimization may thus promote individual foraging specialization (e.g. repeatedly using fast attacks) if the success of a tactic is constant through time. It may also be costly to test different tactics by trial and error when prey are scarce or unpredictable. Alternatively, the costs/risks faced by hunters when switching tactics may be offset through extensive practice, as individuals gain experience and information on their prey, leading to increased individual foraging flexibility [@Stephens1993;@Ishii.Shimada2010;@Mery.Burns2010;@Kelley.Magurran2011;@Snell-Rood2013]. To develop proper responses to fluctuating resources, individuals would need to sample their environment broadly during an extensive period of time, leading to greater performance at later stages of development [reviewed in @Snell-Rood2013]. However, predators can differ in the challenges that they face during their lifetime, which could result in both processes (i.e. specialization and flexibility) operating at the same time within a predator population. For instance, some individuals may hunt in habitats where prey are faster, which are often proven to be more difficult to hunt [@Walker.etal2005;@Kelley.Magurran2011;@Martin.etal2022]. It should thus be essential for predators to learn how to adjust their tactic to their prey to optimize their success.

There is currently a lack of consensus on the fitness advantages of specialized vs flexible foraging strategies [@Phillips.etal2017]. This is reflected in the literature showing contrasting results in the links between specialization and fitness at the individual level. Some studies have found greater success for individuals that use specialized foraging tactics [@Pintor.etal2014;@Patrick.Weimerskirch2014a;@vandenBosch.etal2019] while others found that individuals using flexible foraging tactics have greater success [@Paull.etal2012;@Santoro.etal2019;@Manlick.etal2021]. This disrepancy may in part be due to the limited information we have on the ecological contexts favoring specialization over flexibility. However, many studies have found that both type of foragers can also attain equal success depending on timescales, and suggest that fluctuating resources throughout a predator's lifetime may be key in shaping individual foraging specialization [@Weimerskirch2007;@Woo.etal2008;@Potier.etal2015;@Phillips.etal2017;@Courbin.etal2018]. It is hypothesized that when resources are predictable (or stable), individual specialists should have higher capture rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should benefit from adjusting to fluctuating resources by reducing the consequences of uncertainty [@dall2010]. While it is clear that predators can match their hunting behaviour and locomotion to their prey [@McGhee.etal2013a;@Bro-Jorgensen2013;@Chang.etal2017;@Szopa-Comley.Ioannou2022], the important question remaining is how? Given the dynamic nature of predator-prey interactions, it is unlikely that evolution alone could explain such behavioural adjustments by predators, because complex hunting strategies require practice to be effective [@Kelley.Magurran2011]. Thus, uncovering the role of learning during direct predator-prey interactions would help in better predicting the behavioural decisions and the success of predators when they are hunting.

A recurring challenge impeding research on predator-prey behavioural interactions, at the individual level, is the need to collect data simultaneously on both the predator and prey. We recently showed, using an online predator-prey videogame called *Dead by Daylight* (*DBD*), that virtual systems can overcome this challenge [@FraserFranco.etal2022]. In this dynamic virtual system, four prey players need to forage for resources while avoiding predation by a predator player. The prey can contribute to the group's success by healing or helping others escape the predator, or be selfish and attempt to escape by themselves [@Cere.etal2021]. The predators' hunting tactics and success are driven by the movement of the prey [@FraserFranco.etal2022]. Predators should thus benefit from learning how the prey behave to be successful. Other virtual predator-prey studies show that predation regimes drive individual variation in risk perception [@Beauchamp2020], that familiarity between prey has a positive indirect effect on survival [@Cere.etal2021], and that prey face contrasting natural and social selection regimes (Santostefano et al. in prep). These examples outline how virtual systems could be useful in identifying general ecological patterns, which might in turn help to advance predator-prey research. They also generate large volumes of data on interacting players throughout their lifetime in the game under realistic, controlled, and repeatable ecological scenarios. Hence, virtual systems, including *DBD*, offer the opportunity to tackle fundamental questions about the role of experience on predator-prey interactions.

In this study, we evaluated how hunting experience shapes predator foraging specialization and success using data from players in the online videogame *Dead by Daylight*. First, we investigate how predators develop their individual hunting expertise. We hypothesize that predators should differ in the development of their expertise if they encountered varying levels of difficulty with the prey that they pursued. For example, a predator may face greater difficulty than other individuals if it encountered elusive groups of prey more often. Second, we test the hypothesis that experience will shape individual foraging specialization. We define foraging tactics as the mean movement speed of the predator along a slow-fast continuum, and the level of individual foraging specialization as the intra-individual variance (i.e. IIV) in movement speed from one match to the next, where low IIV translates to individual specialization. If experience reduces the costs of switching between low and high speeds, we predict that individuals (and thus the population) should all converge towards flexible speeds. If experience enables the refinement of the tactics to hunt specific types of prey, then all individuals may instead specialize in their speed. In both scenarios, among individual variation in IIV (i.e. individual specialization) should be low because everyone is either a specialist or a generalist, but the population variance should either increase (flexibility) or decrease (specialization). Alternatively, both processes may operate at the same time if behavioural interactions with the prey vary among predators, such that among individual variation in IIV should increase. In this case, we expect predators that experienced more predictable encounters with their prey (i.e. similar groups across matches) to specialize in their speed, while predators that experienced unpredictable encounters with their prey (i.e. heterogeneous groups across matches) should adopt flexible movement speeds. If we detect such prey-dependent fine-tuning, then all hunters along the flexible-specialist continuum should attain equal success probably because they learned from experience how to adjust their hunting style to their prey.

@@

*DBD* is a survival asymmetric (i.e. a game where the gameplay mechanics differ between two groups) multiplayer online game developed by Behaviour Interactive Inc, in which players can play either as a predator or a prey. Every match includes only one predator and four preys. The objective of the predator is to hunt and capture the prey, and the objective of the prey is to search for resources while avoiding the predator. The resources are in the form of power generators that, once all activated, will enable the prey to escape through one of two exit doors. The composition of the predator and prey group in a match is determined by a skill-based matchmaking algorithm. A match ends when the predator kills all the prey available (i.e. that have not escaped), or when the last remaining prey escapes the virtual environment. Each player, predator or prey, can choose an avatar with abilities that encourage specific play styles (e.g. bold vs cautious prey, or ambush vs roaming predator). During our study period, the game offered `r toString(format(length(unique(dat$avatar_id))))` predator avatars. The virtual environments are composed of fixed and procedurally generated habitat components, such as vegetation, mazes, and buildings. Some of these environments are larger than others, with varying structural complexity. However, predators display only minimal changes in behaviour and hunting success across the environments, probably due to a game feature enabling them to have visual cues of the generators [@FraserFranco.etal2022]. There were `r toString(format(length(unique(dat$environment_id))))` virtual game environments available for play during the study period.

@@

We analyzed the predator's average speed and the average speed of the prey group encountered by the predator. The predator's average speed is measured as the average distance traveled per second during a match (mean= `r toString(format(round(mean(dat$pred_speed), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$pred_speed), digits = 2), nsmall = 2))` m/s). We measured the preys' average speed as the average travel speed of the four individual prey within a match (mean = `r toString(format(round(mean(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` m/s). We defined hunting success as the number of prey consumed during the match (min = `r toString(format(min(dat$hunting_success)))`, max = `r toString(format(max(dat$hunting_success)))`). Lastly, we defined the predator's cumulative experience as the number of matches played prior to the match being monitored. For example, the first match of a player would have a cumulative experience value of 0, while the tenth match would have a value of 9. We also categorized predators for each match based on their cumulated experience. We labeled predators as novices for matches where they had cumulated less than 100 matches, intermediate for matches where they had cumulated between 100 and 299 matches, and advanced for matches where they had cumulated more than 299 matches (max 499). Since our goal was to monitor predator players throughout their experience and that they all played at least 300 matches, they appeared in all the three experience categories.

@@

The models were fitted in R (version 4.1.2) using Markov chain Monte Carlo (MCMC) sampling with the package "brms" version 2.16.3 [@Burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@Gabry.Cesnovar2021] as the back-end for estimation (cmdstan installation version 2.28.2). All our statistical analyses were executed on Cedar (https://docs.alliancecan.ca/wiki/Cedar), a computer cluster maintained by the Digital Research Alliance of Canada. The operating system for Cedar is CentOS Linux 7.

@@

We tested how prey influence the development of predator expertise by computing five Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success (i.e. number of prey consumed) and the predators' cumulative experience (i.e. number of matches played before the current match). We parameterized the models following the method of @Pedersen.etal2019. The first model was the simplest, with a common global smoothing function and random intercepts for the predator ID. In this model, we assume that predators have the same development of expertise, with the model estimating a trend for the average individual (i.e. global smoother). For the second model, we included varying group-level smoothers for the predator ID. Here, we assume that individual predators share a similar relationship between success and experience, but that this relationship can vary among them (e.g. predator 1 has a steeper curve than predator 2). This enabled us to test whether predators differed in the development of their expertise. In the third model, we kept the group-level smoothers for the individual predators, but removed the global smoother. This model assumes that predators do not share a common relationship between success and experience. The fourth and fifth models were a reproduction of the second and third models respectively, where we included the prey group's average speed to assess its effect on the relationship between success and experience. We included the standardized match duration and average prey rank (a proxy for prey skill) as covariates in all five models. The ranking system in *DBD* was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank), and failing to account for it would prevent us from detecting a change in the predator's foraging success with experience.

@@

The model simultaneously estimated three distinct components for each experience level and each trait. The first component is among individual variation in mean speed and success (the mean part of the model), allowing to test for individual differences in tactic use (i.e. along the slow-fast continuum), in the mean speed of the prey encountered, and in prey consumption. The second component is among individual variation in intra-individual behavioural variation (IIV) (the dispersion part of the model), allowing us to test for individual differences in foraging specialization and in how (dis)similar was the speed of the groups of prey encountered (i.e. if resource heterogeneity was higher for some individuals compared to others). The third component consists of pairwise comparisons between the individual identity random effect variances on the mean and dispersion part of the model across traits and experience levels, computed as a 15x15 (co)variance matrix. This allows us to assess whether predators that are on average slower/faster are more specialized or flexible, whether predators that adopted a slower/faster (or more specialized/flexible) tactic encountered prey that were faster/slower, and whether slower/faster predators had lower/greater hunting success.

We controlled for the standardized average rank of the prey group for both the predator and the prey's speed on both the mean and the dispersion parts of the model. We included random effects for the game environment, the predator avatar, and the individual predator identity on the mean part of the model, and an individual identity random effect exclusively on the dispersion part. For hunting success, we modeled the mean part only and controlled for the standardized mean prey rank and game duration, and included an individual identity random effect.

@@

where $j$ is the index of the experience levels. The model is fitted on $n$ experience levels ($n = 3$) denoted by $T_{j}$, where $j$ ranges from 1 to $n$, being the novice, intermediate, and advanced levels. The $b$ terms are vectors of fixed effects with their incidence matrices $X$. The terms $en$, $av$, and $id$ are the vectors of random environment, random avatar, and random individual identity effects associated with their incidence matrices $M$, $N$, and $Z$. The $d$ subscript indicates that a term is from the dispersion part of the model. On the mean part for all traits and experience levels, we assumed that the game environment and avatar random effects followed a Gaussian distribution with estimated standard deviation ($en \sim N(0,\ I\ \sigma_{en})$ and $av \sim N(0,\ I\ \sigma_{av})$). $I$ is the identity matrix, with $I\sigma$ indicating that the random effects are independently and identically distributed. The residuals are assumed to follow a Gaussian distribution ($e \sim N(0,\ Diag \{\sigma_{e}\})$) *explain here that each has their own individual variance*. The individual identity random effects on both the mean and dispersion parts across all response variables and experience levels are indexed in vector $u$. The vector $u$ of individual identity random effects follows a multivariate Gaussian distribution $u \sim MVN(0,\ \Sigma)$ where $\Sigma$ is a 15x15 (co)variance matrix.

@@

We parametrized both the GAMMs and the MDHMLM to run four MCMC chains with 1000 posterior samples for each parameter. We ran 2500 iterations with a thinning set to eight for the additive model with a global smoother only (see Table 1), and 1500 iterations with a thinning set to four for the other additive models. We set the first 500 iterations in each model as warm ups. We parametrized the MDHGLM to run 2500 iterations with a thinning set to eight, with the first 500 iterations used as warm ups. We assessed the convergence of the MCMC chains using trace plots, R-hat diagnostics with a threshold of <1.01, and effective sample sizes (ESS) with a threshold of >100 [@Vehtari.etal2021a]. We also performed posterior predictive checks which showed an adequate fit of the models (for details, see https://github.com/quantitative-ecologist/experience-hunting-tactics). We report all the posterior parameter estimates using the median of the posterior distribution (or the mean if stated otherwise) with the highest posterior density (HPD) intervals at 50%, 80%, and 95%.

>>>>

Individual variation in predator foraging behaviour is increasingly recognized as a major driver of trophic interactions and community dynamics [@Griffen.etal2012;@Michalko.Pekar2016;@Moran.etal2017;@Michalko.etal2021]. Indeed, predator populations often comprise assemblages of individuals specializing on different foraging tactics, with consequences for habitat use, functional responses, prey choice, and foraging rate [@Kobler.etal2009;@toscanoTraitmediatedFunctionalResponses2014;@Patrick.Weimerskirch2014;@Matsumura.Miyatake2022a]. Despite clear evidence that individual specialization is important in predator-prey interactions, its underlying mechanisms remain surprisingly unclear. One hypothesis suggests that reciprocal interactions between predators and prey could drive individual foraging specialization [@Araujo.etal2011;@toscanoPersonalityForagingBehavior2016;@DeLisle.etal2022]. For instance, predators often hunt using techniques that are fine-tuned to the type of prey that they encounter [@Estes.etal2003;@Woo.etal2008;@Courbin.etal2018], and their refinement is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@Heithaus.etal2018]. Thus, the capacity of predators to effectively use hunting tactics should be contingent on periods of extensive practice (i.e. experience), which may result in different foraging specializations.

Theory predicts that foraging specialization may emerge via learning and expertise acquisition [@Tinker.etal2009;@Dukas2019]. @Dukas2019 defines expertise as the characteristics, skills, and knowledge that allow individuals with extensive experience to outperform novices on complex tasks. It is described as the relationship between success and experience, where extensive practice optimizes success when individuals reach an advanced level of experience. Empirical studies on human and non-human hunters show that experience optimizes foraging efficiency (e.g. search and handling times, return rates) potentially via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@MacDonald2007;@Reid.etal2010;@Wilson-Rankin2015]. This body of work suggests that such optimization may thus promote individual foraging specialization (e.g. repeatedly using fast attacks) if the success of a tactic is constant through time. It may also be costly to test different tactics by trial and error when prey are scarce or unpredictable. Alternatively, the costs/risks faced by hunters when switching tactics may be offset through extensive practice, as individuals gain experience and information on their prey, leading to increased individual foraging flexibility [@Stephens1993;@Ishii.Shimada2010;@Mery.Burns2010;@Kelley.Magurran2011;@Snell-Rood2013]. To develop proper responses to fluctuating resources, individuals would need to sample their environment broadly during an extensive period, leading to greater performance at later stages of development [reviewed in @Snell-Rood2013]. However, predators can differ in the challenges that they face during their lifetime, which could result in both processes (i.e. specialization and flexibility) operating at the same time within a predator population. For example, some individuals may hunt in habitats where prey are faster, which are often proven to be more difficult to hunt [@Walker.etal2005;@Kelley.Magurran2011;@Martin.etal2022]. Thus, it should be essential for predators to learn how to adjust their tactic to their prey to optimize their success.

There is currently a lack of consensus on the fitness advantages of specialized vs flexible foraging strategies [@Phillips.etal2017]. This is reflected in the literature showing contrasting results in the links between specialization and fitness at the individual level. Some studies have found greater success for individuals that use specialized foraging tactics [@Pintor.etal2014;@Patrick.Weimerskirch2014a;@vandenBosch.etal2019] while others found that individuals using flexible foraging tactics have greater success [@Paull.etal2012;@Santoro.etal2019;@Manlick.etal2021]. This disrepancy may in part be due to the limited information we have on the ecological contexts favoring individual specialization over flexibility. However, many studies have found that both type of foragers can also attain equal success depending on timescales, and suggest that fluctuating resources throughout a predator's lifetime may be key in shaping individual foraging specialization [@Weimerskirch2007;@Woo.etal2008;@Potier.etal2015;@Phillips.etal2017;@Courbin.etal2018]. It is hypothesized that when resources are predictable (or stable), individual specialists should have higher capture rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should benefit from adjusting to fluctuating resources by reducing the consequences of uncertainty [@dall2010]. While it is clear that predators can match their hunting behaviour and locomotion to their prey [@McGhee.etal2013a;@Bro-Jorgensen2013;@Chang.etal2017;@Szopa-Comley.Ioannou2022], the important question remaining is how? Given the dynamic nature of predator-prey interactions, it is unlikely that evolution alone could explain such behavioural adjustments by predators, because complex hunting strategies require practice to be effective [@Kelley.Magurran2011]. Thus, uncovering the role of learning during direct predator-prey interactions would help to predict the behavioural decisions and the success of predators when they are hunting.

A recurring challenge impeding research on predator-prey behavioural interactions, at the individual level, is the need to collect data simultaneously on both the predator and the prey. We recently showed, using an online predator-prey videogame called *Dead by Daylight* (*DBD*), that virtual systems can overcome this challenge [@FraserFranco.etal2022]. In this dynamic virtual system, four prey players need to forage for resources while avoiding predation by a fifth player. The prey can contribute to the group's success by healing or helping others escape from the predator, or be selfish and attempt to escape by themselves [@Cere.etal2021]. The predators' hunting tactics and success are driven by the movement of the prey [@FraserFranco.etal2022]. Predators should thus benefit from learning how the prey behave to be successful. Other virtual predator-prey studies show that predation regimes drive individual variation in risk perception [@Beauchamp2020], that familiarity between prey has a positive indirect effect on survival [@Cere.etal2021], and that prey face contrasting natural and social selection regimes (Santostefano et al. in prep). These examples outline how virtual systems could be useful in identifying general ecological patterns, which in turn might help to advance predator-prey research. They also generate large volumes of data on interacting players throughout their lifetime in the game under realistic, controlled, and repeatable ecological scenarios. Hence, virtual systems, including *DBD*, offer the opportunity to tackle fundamental questions about the role of experience on predator-prey interactions.

In this study, we evaluated how hunting experience shapes predator foraging specialization and success using data from players in the online videogame *Dead by Daylight*. First, we investigate how predators develop their individual hunting expertise. We hypothesize that predators should differ in the development of their expertise if they encountered varying levels of difficulty with the prey that they pursued. For example, a predator may face greater difficulty than other individuals if it encountered elusive groups of prey more often. Second, we test the hypothesis that experience will shape individual foraging specialization. We define foraging tactics as the mean movement speed of the predator along a slow-fast continuum, and the level of individual foraging specialization as the intra-individual variance (i.e. IIV) in movement speed across matches, where low IIV translates to individual specialization (e.g. always using a fast tactic). If experience reduces the costs of switching between low and high speeds as predators encounter varying groups of prey, we predict that individuals (and thus the population) should all converge towards flexible speeds. If experience increases a tactic's efficiency for specific types of prey assuming that all individuals encounter similar groups of prey, then they may instead specialize in their speed. In both scenarios, among individual variation in IIV (i.e. individual differences in specialization) should be low across experience because everyone is either a flexible or a specialist forager, but the population variance would either increase (flexibility) or decrease (specialization). Alternatively, both processes may operate at the same time if behavioural interactions with the prey vary among predators, such that among individual variation in IIV should increase. In this case, we expect predators that experienced more predictable encounters with their prey (i.e. similar groups across matches) to specialize in their speed, while predators that experienced unpredictable encounters with their prey (i.e. heterogeneous groups across matches) should adopt flexible movement speeds. If we detect such prey-dependent fine-tuning, then all hunters along the flexible-specialist continuum should attain equal success probably because they learned from experience how to adjust their hunting style to their prey.

@@

*DBD* is a survival asymmetric multiplayer online game (i.e. a game where the gameplay mechanics differ between two groups) developed by Behaviour Interactive Inc, in which players can play either as a predator or a prey. Every match includes only one predator and four prey. The objective of the predator is to hunt and capture the prey, and the objective of the prey is to search for resources while avoiding the predator. The resources are in the form of power generators that, once all activated, will enable the prey to escape through one of two exit doors. The composition of the predator and prey group in a match is determined by a skill-based matchmaking algorithm. A match ends when the predator kills all the prey available (i.e. that have not escaped), or when the last remaining prey escapes the virtual environment. Each player, predator or prey, can choose an avatar with abilities that encourage specific play styles (e.g. bold vs cautious prey, or ambush vs roaming predator). During our study period, the game offered `r toString(format(length(unique(dat$avatar_id))))` predator avatars. The virtual environments are composed of fixed and procedurally generated habitat components, such as vegetation, mazes, and buildings. Some of these environments are larger than others, with varying structural complexity. However, predators display only minimal changes in behaviour and hunting success across the environments, probably due to a game feature enabling them to have visual cues of the generators [@FraserFranco.etal2022]. There were `r toString(format(length(unique(dat$environment_id))))` virtual game environments available for play during the study period.

@@

We analyzed the predator's average speed and the average speed of the prey group encountered by the predator. The predator's average speed is measured as the average distance traveled per second during a match (mean= `r toString(format(round(mean(dat$pred_speed), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$pred_speed), digits = 2), nsmall = 2))` m/s). We measured the preys' average speed as the average travel speed of the four individual prey within a match (mean = `r toString(format(round(mean(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` ± `r toString(format(round(sd(dat$prey_avg_speed, na.rm = T), digits = 2), nsmall = 2))` m/s). We defined hunting success as the number of prey consumed during the match (min = `r toString(format(min(dat$hunting_success)))`, max = `r toString(format(max(dat$hunting_success)))`). Lastly, we defined the predator's cumulative experience as the number of matches played prior to the match being monitored. For example, the first match of a player would have a cumulative experience value of 0, while the tenth match would have a value of 9. We also categorized predators for each match based on their cumulated experience. We labeled predators as novices for matches where they had cumulated less than 100 matches, intermediate for matches where they had cumulated between 100 and 299 matches, and advanced for matches where they had cumulated more than 299 matches (max 499). Since our goal was to monitor predator players throughout their experience and that they all played at least 300 matches, they all appeared in each of the three experience categories.

@@

All models were fitted under a Bayesian framework in R (version 4.1.2) using Markov chain Monte Carlo (MCMC) sampling with the package "brms" version 2.16.3 [@Burkner2017], an R front-end for the STAN software [@StanDevelopmentTeam2023], and "cmdstanr" version 0.4.0 [@Gabry.Cesnovar2021] as the back-end for estimation (cmdstan installation version 2.28.2). The models were compiled on Cedar, a computer cluster maintained by the Digital Research Alliance of Canada (https://docs.alliancecan.ca/wiki/Cedar). The operating system for Cedar is CentOS Linux 7, and each model required 64GB of RAM with 48 cores to compile within $\le$5 days.

@@

We tested how prey influenced the development of predator expertise by computing five Bayesian generalized additive mixed models (GAMM) with thin-plate regression splines, all of which estimated the relationship between hunting success (i.e. number of prey consumed) and the predators' cumulative experience (i.e. number of matches played before the current match). We parameterized the models following the method of @Pedersen.etal2019. The first model was the simplest, with a common global smoothing function and random intercepts for the predator ID. In this model, we assume that predators have the same development of expertise, with the model estimating a trend for the average individual (i.e. global smoother). For the second model, we included varying individual smoothers for the predator ID. Here, we assume that individual predators share a similar relationship between success and experience, but that this relationship can vary among them (e.g. predator 1 has a steeper curve than predator 2). This enabled us to test whether predators differed in the development of their expertise. In the third model, we kept the individual smoothers for the predators, but removed the global smoother. This model assumes that predators do not share a common relationship between success and experience. The fourth and fifth models were a reproduction of the second and third models respectively, where we included the prey group's average speed to assess its effect on the relationship between success and experience. We included the standardized match duration and average prey rank (a proxy for prey skill) as covariates in all five models. The ranking system in *DBD* was implemented by the company to pair players in a match based on their skill (https://deadbydaylight.fandom.com/wiki/Rank), and failing to account for it would prevent us from detecting a change in the predator's foraging success with experience.

@@

The model simultaneously estimated three distinct components for each experience level and each trait. The first component is among individual variation in mean speed and success (the mean part of the model), allowing to test for individual differences in tactic use (i.e. along the slow-fast continuum), in the mean speed of the prey encountered, and in mean prey consumption. The second component is among individual variation in intra-individual behavioural variation (IIV) (the dispersion part of the model), allowing us to test for individual differences in foraging specialization and in how (dis)similar were the groups of prey encountered (i.e. if resource heterogeneity was higher for some individuals compared to others). The third component consists of pairwise comparisons between the individual identity random effect standard deviations on the mean and dispersion part of the model within/across traits within experience levels, and within/across traits across experience levels, resulting in a 15x15 (co)variance matrix. This allows us to assess whether predators that are on average slower/faster are more specialized or flexible, whether predators that adopted a slower/faster (or more specialized/flexible) tactic encountered prey that were faster/slower, and whether slower/faster predators had lower/greater hunting success.

We controlled for the standardized average rank of the prey group for both the predator and the prey's speed on both the mean and the dispersion parts of the model. We included game environment and predator avatar random effects on the mean part of the model exclusively that were trait and experience specific (i.e. no covariance structure). We included individual predator identity random effects on the mean and dispersion parts. For hunting success, we modeled the mean part only and controlled for the standardized mean prey rank and game duration, and included an individual identity random effect.

@@

where $j$ is the index of the experience levels. The model is fitted on $n$ experience levels ($n = 3$) denoted by $T_{j}$, where $j$ ranges from 1 to $n$, being the novice, intermediate, and advanced levels. The $b$ terms are vectors of fixed effects with their design matrices $X$. The terms $en$, $av$, and $id$ are the vectors of random environment, random avatar, and random individual identity effects associated with their design matrices $M$, $N$, and $Z$. The $d$ subscript indicates that a term is from the dispersion part of the model. On the mean part for all traits and experience levels, we assumed that the game environment and avatar random effects followed a Gaussian distribution with estimated standard deviation ($en \sim N(0,\ I\ \sigma_{en})$ and $av \sim N(0,\ I\ \sigma_{av})$). $I$ is the identity matrix, with $I\sigma$ indicating that the random effects are independently and identically distributed. The individual identity random effects on both the mean and dispersion parts across all response variables and experience levels are indexed in vector $u$. The individual identity random effects contained in vector $u$ follow a multivariate Gaussian distribution $u \sim MVN(0,\ \Sigma)$ where $\Sigma$ is a 15x15 (co)variance matrix of the individual identity standard deviations $\sigma_{id_i}$ across all experience levels and response variables.

@@

We parametrized both the GAMMs and the MDHMLM to run four MCMC chains with 1000 posterior samples for each parameter. We ran 2500 iterations with a thinning set to eight for the additive model with a global smoother only (see Table 1), and 1500 iterations with a thinning set to four for the other additive models. We set the first 500 iterations in each model as warm ups. We parametrized the MDHGLM to run 2500 iterations with a thinning set to eight, with the first 500 iterations used as warm ups. We assessed the convergence of the MCMC chains using trace plots, R-hat diagnostics with a threshold of <1.01, and effective sample sizes (ESS) with a threshold of >100 [@Vehtari.etal2021a]. We also performed posterior predictive checks which showed an adequate fit of the models (for details, see https://github.com/quantitative-ecologist/experience-hunting-tactics). We report all the posterior parameter estimates using the median of the posterior distribution with the highest posterior density (HPD) intervals at 50%, 80%, and 95%.
